
Note: Due to confidentiality some modifications have been done to the data.
Areas with less than 10 observed cases have been modificated.

## Read the data

```{r}
Data=read.table(file="DataExample.txt")
```

## Read the R-INLA package

```{r}
library(INLA)
```

## Description of indexes created to fit the final model

1)  i_P: take the value of 1 if the observation corresponds to lung cancer and 0 if the observation corresponds to LOCP cancer.

2)  i_LOCP: take the value of 1 if the observation corresponds to LOCP cancer and 0 if the observation corresponds to lung cancer.

3)  i: a region index. Takes values 1 to 94. Note the model="besag2" defines a model with length $N=2n$ where $n$ is the size of the graph. In this case $2*47$.

4)  LOCP_het: take the value of the Region number (1 to 47) if the observation corresponds to LOCP cancer and NA if the observation corresponds to lung cancer.

## Read the graph using a neighbourhood matrix

```{r}
g <- inla.read.graph("esp_prov_nb.inla")
```

The .inla file contains the following:

47\
1 5 8 19 25 30 45\
2 7 3 12 15 17 22 29 43\
3 3 2 29 43\
4 2 17 29\
$\vdots$\
42 6 5 6 9 12 15 27\
43 5 2 3 11 15 41\
44 7 5 8 23 33 35 37 46\
45 4 1 8 19 36\
46 4 23 31 35 44\
47 8 18 21 24 25 30 39 40 41\

The first line summarized the total number of regions, i.e. 47.

The second line indicates that the province 1 has 5 neighbours that are provinces 8, 19, 25, 30, 45.

The third line indicates that the province 2 has 7 neighbours that are provinces 3, 12, 15, 17, 22, 29, 43.

And so on.

## PC-priors definition

```{r}
pcprec <- list(theta=list(prior='pc.prec', param=c(1,0.01)))
```

## Formula for the final model and INLA execution

```{r}
formula = Cases~ -1 + i_P + i_LOCP + f(i, model="besag2", graph=g, hyper=pcprec, replicate=Gender, scale.model = T)   +
  f(LOCP_het, model="iid", hyper=pcprec) +
  f(Age_group, model="iid", hyper=pcprec, replicate=Gender)

```

```{r}
result = inla(formula, family = "poisson",
              data = data.frame(Data), E=Population,
              control.compute=list(dic=TRUE, cpo=TRUE, waic=TRUE, hyperpar=TRUE),
              control.inla=list(strategy="simplified.laplace",verbose = F, numint.maxfeval= 100000))
```

## Selection criteria values for the model

### DIC

```{r}
result$dic$dic    
```

### WAIC

```{r}
result$waic$waic
```

### LS

```{r}
logscore=sum(-log(result$cpo$cpo)) 
logscore
```

## Results

Require packages:

```{r}
library(RColorBrewer)
library(maptools)
library(Hmisc)
library(R2WinBUGS)
library(spdep)
library(sp)
library(Rgraphviz)
library(classInt)    
library(lattice)
```

Load Cartography of continental Spain :

```{r}
xName<- readShapePoly("ZonificacionEspanaProv/esp_prov.shp",IDvar="ESP_PROV_I", proj4string=CRS("+proj=longlat +ellps=clrk66"))

xSC <- xName[xName$ESP_PROV_I !=35 &xName$ESP_PROV_I !=38 &xName$ESP_PROV_I !=51 &
               xName$ESP_PROV_I !=52 & xName$ESP_PROV_I !=7,]
```

### Figure 4: Posterior medians of the rates for age-groups 30-44 and 45-59

Calculation of the posterior medians and the 95% confident interval of the rates

```{r}
tasae<-result$summary.fitted.values[,c(3:5)]
colnames(tasae)=c("ilimt","tasae","isupt")
```

We create a new data frame called AreaAgeGender, with the information of the gender, age-group and province, the population of each level and the estimated posterior medians and confident intervals. We want to plot the estimated posterior median of the rates per 100,000 population.

```{r}
AreaAgeGender<-data.frame(cbind(Data[,c(1:4,6)],tasae))

AreaAgeGender$tasae=AreaAgeGender$tasae*10^5
tail(AreaAgeGender)
```

We select the rates of 30-44 and 45-59 age-groups and we added to the cartography.

```{r}
D<-length(unique(AreaAgeGender$Cancer))
disease <- c("LUNG","LOCP")

G<-length(unique(AreaAgeGender$Gender))
gender <- c("M","F")
J<-length(unique(AreaAgeGender$Age_group))

comarc=c(unique(Data$Region_n))

for (i in 1:D) {
  for (k in 1:G) {
    for (l in 1:2) {
      eval(parse(text=paste0("pr",disease[i],l,gender[k],"<-AreaAgeGender[AreaAgeGender$Cancer==",i," & AreaAgeGender$Gender==",k," & AreaAgeGender$Age_group==",l,",]$tasae")))
      
      eval(parse(text=paste0("region",disease[i],l,gender[k],"<-data.frame(comarc,pr",disease[i],l,gender[k],")")))
      eval(parse(text=paste0("colnames(region",disease[i],l,gender[k],") <- c('region','crude_rate')")))
      eval(parse(text=paste0("xSC$rate",disease[i],l,gender[k],"<-region",disease[i],l,gender[k],"[,2]")))
    }
  }
}
```

Lung cancer mortality rates for 30-44 age-group by regions and gender:

```{r}
agelabels=c( "30-44", "45-59", "60-74", "75-89", "90+")

colorkeypval=list(labels=paste(seq(1.2,4.2,by=0.3)),
                  at=seq(1.2,4.2,by=0.3)) 

spplot(xSC,zcol=c("rateLUNG1M","rateLUNG1F"),
       names.attr=c("male","female"),at=seq(1.2,4.2,by=0.3),
       as.table=T,col.regions=c("#ffffff","#FFFFCC","#FFEDA0","#FED976","#FEB24C","#FD8D3C","#FC4E2A","#E31A1C","#BD0026","#800026"),
       colorkey=colorkeypval)
```

Lung cancer mortality rates for 45-59 age-group by regions and gender:

```{r}
colorkeypval=list(labels=paste(c(seq(12,28, by=4),seq(35.5,80.5,by=7.5))),
                  at=c(seq(12,28, by=4),seq(35.5,80.5,by=7.5))) 

spplot(xSC,zcol=c("rateLUNG2M","rateLUNG2F"),
       names.attr=c("male","female"),at=c(seq(12,28, by=4),seq(35.5,80.5,by=7.5)),
       as.table=T,col.regions=c("#ffffff","#FFFFCC","#FFEDA0","#FED976","#FEB24C","#FD8D3C","#FC4E2A","#E31A1C","#BD0026","#800026","black"),
       colorkey=colorkeypval)
```

LOCP cancer mortality rates for 30-44 age-group by regions and gender:

```{r}
colorkeypval=list(labels=paste(seq(0.1,0.5,by=0.05)),
                  at=seq(0.1,0.5,by=0.05)) 

spplot(xSC,zcol=c("rateLOCP1M","rateLOCP1F"),
       names.attr=c("male","female"),at=seq(0.1,0.5,by=0.05),
       as.table=T,col.regions=c("#ffffff","#FFFFCC","#FFEDA0","#FED976","#FEB24C","#FD8D3C","#FC4E2A","#E31A1C","#BD0026","#800026"),
       colorkey=colorkeypval)

```

LOCP cancer mortality rates for 45-59 age-group by regions and gender:

```{r}
colorkeypval=list(labels=paste(seq(1.5,9,by=0.75)),
                  at=seq(1.5,9,by=0.75)) 

spplot(xSC,zcol=c("rateLOCP2M","rateLOCP2F"),
       names.attr=c("male","female"),at=seq(1.5,9,by=0.75),
       as.table=T,col.regions=c("#ffffff","#FFFFCC","#FFEDA0","#FED976","#FEB24C","#FD8D3C","#FC4E2A","#E31A1C","#BD0026","#800026"),
       colorkey=colorkeypval)
```

### Figure 5: Posterior medians of the rates for 60-74, 75-89 and 90+ age-groups

We select the rates of 60-74, 75-89 and 90+ age-groups and we added to the cartography

```{r}
for (i in 1:D) {
  for (k in 1:G) {
    for (l in 3:J) {
      eval(parse(text=paste0("pr",disease[i],l,gender[k],"<-AreaAgeGender[AreaAgeGender$Cancer==",i," & AreaAgeGender$Gender==",k," & AreaAgeGender$Age_group==",l,",]$tasae")))
      
      eval(parse(text=paste0("region",disease[i],l,gender[k],"<-data.frame(comarc,pr",disease[i],l,gender[k],")")))
      eval(parse(text=paste0("colnames(region",disease[i],l,gender[k],") <- c('region','crude_rate')")))
      eval(parse(text=paste0("xSC$rate",disease[i],l,gender[k],"<-region",disease[i],l,gender[k],"[,2]")))
    }
  }
}
```

Lung cancer mortality rates for males by regions and age-groups

```{r}
colorkeypval=list(labels=paste(seq(165,520,by=35)),
                  at=seq(165,520,by=35))
spplot(xSC,zcol=c("rateLUNG3M","rateLUNG4M","rateLUNG5M"),
       names.attr=c("60-74","75-89","90+"),at=seq(165,520,by=35),
       as.table=T,col.regions=c("#F7FBFF","#DEEBF7","#C6DBEF","#9ECAE1","#6BAED6","#4292C6","#2171B5","#08519C","#08306B","black"),
       colorkey=colorkeypval)
```

Lung cancer mortality rates for females by regions and age-groups

```{r}
colorkeypval=list(labels=paste(seq(20,105,by=8.5)),
                  at=seq(20,105,by=8.5)) 

spplot(xSC,zcol=c("rateLUNG3F","rateLUNG4F","rateLUNG5F"),
       names.attr=c("60-74","75-89","90+"),at=seq(20,105,by=8.5),
       as.table=T,col.regions=c("#F7FBFF","#DEEBF7","#C6DBEF","#9ECAE1","#6BAED6","#4292C6","#2171B5","#08519C","#08306B","black"),
       colorkey=colorkeypval)
```

LOCP cancer mortality rates for males by regions and age-groups

```{r}
colorkeypval=list(labels=paste(seq(15,60,by=5)),
                  at=seq(15,60,by=5)) 

spplot(xSC,zcol=c("rateLOCP3M","rateLOCP4M","rateLOCP5M"),
       names.attr=c("60-74","75-89","90+"),at=seq(15,60,by=5),
       as.table=T,col.regions=c("#F7FBFF","#DEEBF7","#C6DBEF","#9ECAE1","#6BAED6","#4292C6","#2171B5","#08519C","#08306B","black"),
       colorkey=colorkeypval)
```

LOCP cancer mortality rates for females by regions and age-groups

```{r}
colorkeypval=list(labels=paste(seq(3,18,by=1.5)),
                  at=seq(3,18,by=1.5)) 

spplot(xSC,zcol=c("rateLOCP3F","rateLOCP4F","rateLOCP5F"),
       names.attr=c("60-74","75-89","90+"),at=seq(3,18,by=1.5),
       as.table=T,col.regions=c("#F7FBFF","#DEEBF7","#C6DBEF","#9ECAE1","#6BAED6","#4292C6","#2171B5","#08519C","#08306B","black"),
       colorkey=colorkeypval)
```

### Table 1: Ranking

We want to see the estimated posterior median and the 95% credible interval of the rates per 100,000 population for the 10 provinces with highest mortality rates.

```{r}
disease <- c("LUNG","LOCP")
gender <- c("Male","Female")

AreaAgeGender$isupt<-AreaAgeGender$isupt*10^5
AreaAgeGender$ilimt<-AreaAgeGender$ilimt*10^5

for (i in 1:D) {
  for (k in 1:G) {
    for (l in 1:J) {
      eval(parse(text=paste0(disease[i],"_",gender[k],"_",l,"<- AreaAgeGender[which(AreaAgeGender$Cancer==i & AreaAgeGender$Gender==k & AreaAgeGender$Age_group==l),]")))
      
      eval(parse(text=paste0( disease[i],"_",gender[k],"_",l,"<-",disease[i],"_",gender[k],"_",l,"[order(",disease[i],"_",gender[k],"_",l,"$tasae, decreasing=TRUE),]")))
    }
    
  }
}
```

Top 10 provinces for 60-74, 75-89 and 90+ age-groups by gender

```{r}
for (i in 1:D) {
  for (k in 1:G) {
    for (l in 3:J) {
      eval(parse(text=paste0("table<-",disease[i],"_",gender[k],"_",l,"[c(1:10),c(3,6:8)]")))
      latex_table<-xtable::xtable(table, caption=(text=paste0("Top 10 provinces for ",gender[k]," cancer type",disease[i]," and age-group",l)), digits=3)
      xtable::print.xtable(latex_table, include.rownames = FALSE, comment=FALSE, caption.placement = getOption("xtable.caption.placement", "top"))
    }
  }
}

```

### Figure 6: Posterior gender-specific shared spatial pattern

The results of the model gives the spatial pattern for each disease and gender, that is to say, $\delta\kappa_{ig}$ for lung cancer and $\frac{1}{\delta}\kappa_{ig}$ for LOCP cancer. So we get the posterior median of the parameter $\delta$, to obtain the posterior median and credible interval of $\kappa_{ig}$.

```{r}
delta <-result$summary.hyperpar$`0.5quant`[2]

reg_index=188
csi <- result$marginals.random$i[1:reg_index] 

zeta <- lapply(csi,function(x) inla.tmarginal(exp,x))

spatialeffect=NULL
for(l in 1:max(unique(reg_index))){
  spatialeffect=rbind(spatialeffect,c(inla.qmarginal(0.5, zeta[[l]]),inla.hpdmarginal(0.95,zeta[[l]])))
}#median, low-limit, upper-limit.

```

```{r}
xSC$LUNGM <- spatialeffect[1:47]/delta
xSC$LUNGF <- spatialeffect[95:141]/delta
xSC$LOCPM <- spatialeffect[48:94]*delta
xSC$LOCPF <- spatialeffect[142:188]*delta
```

We calculate the median for each gender by provinces. We calculate the median because we have the posterior distribution of $\delta\kappa_{ig}$ and $\frac{1}{\delta}\kappa_{ig}$, and when we undo the effect of $\delta$, we can do a little mistake because we use only the median of the distribution.

```{r}
xSC$MALES <- (xSC$LUNGM+xSC$LOCPM)/2

xSC$FEMALES <- (xSC$LUNGF+xSC$LOCPF)/2
```

We plot the posterior gender-specific shared spatial pattern

```{r}
colorkeypval=list(labels=paste(seq(0.3,2.5,by=0.2)),
                  at=seq(0.3,2.5,by=0.2)) 

spplot(xSC,zcol=c("MALES","FEMALES"),
       names.attr=c("Males","Females"),at=seq(0.3,2.5,by=0.2),
       as.table=T,col.regions=c("#ffffff",'#ffffe5','#fff7bc','#fee391','#fec44f','#fe9929','#ec7014','#cc4c02','#993404','#662506',"black"),
       colorkey=colorkeypval)

```

### Figure 7: Posterior median of the spatially unstructures effect

The results of the model gives the posterior distribution of the spatially unstructured random effect of LOCP cancer. We get the posterior median and credible interval of the effect $u_i$ and we plot the posterior median of the spatially unstructured random effect pattern for LOCP cancer mortality.

```{r}
csi <-result$marginals.random$LOCP_het
zeta <- lapply(csi,function(x) inla.tmarginal(exp,x))

spatialunstructures=NULL
for(l in 1:47){
  spatialunstructures=rbind(spatialunstructures,c(inla.qmarginal(0.5, zeta[[l]]),inla.hpdmarginal(0.95,zeta[[l]])))
}#median, low-limit, upper-limit.

xSC$LOCPuns <- spatialunstructures[1:47,1]

colorkeypval=list(labels=paste(seq(0.75,1.65,by=0.1)),
                  at=seq(0.75,1.65,by=0.1)) 

spplot(xSC,zcol=c("LOCPuns"),
       names.attr=c("LOCP"),at=seq(0.75,1.65,by=0.1),
       as.table=F,col.regions=c("#ffffff","#FFFFCC","#FFEDA0","#FED976","#FEB24C","#FD8D3C","#FC4E2A","#E31A1C","#BD0026","#800026"),
       colorkey=colorkeypval)
```
